---
title: "Michael_Ghattas_WP8"
author: "Michael Ghattas"
date: "`r Sys.Date()`"
output: pdf_document
---

# START:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load necessary libraries
library(tidyverse)
library(glmnet)
library(MASS)
library(AER)
```

```{r}
# Load data
data(NMES1988)
NMES1988 <- NMES1988 %>% mutate(emergency = visits)

# Split the data into training and test sets
set.seed(1)
n <- nrow(NMES1988)
train_index <- sample(1:n, size = 0.7 * n)
train <- NMES1988[train_index, ]
test <- NMES1988[-train_index, ]
```

```{r}
# Fit a Negative Binomial model for emergency room visits
negbin_model <- glm.nb(emergency ~ ., data = train)
summary(negbin_model)

# Display the Negative Binomial model summary
summary(negbin_model)
```

```{r}
# Variable selection using Lasso
X_train <- model.matrix(emergency ~ . - 1, data = train)
Y_train <- train$emergency

# Fit Lasso using cross-validation
set.seed(3456787)
lasso_model <- cv.glmnet(X_train, Y_train, alpha = 1, family = "poisson")
lasso_lambda_min <- lasso_model$lambda.min
coef(lasso_model, s = lasso_lambda_min)
```

```{r}
# Cross-validation plot for Lasso
plot(lasso_model)
title("Lasso Cross-Validation Plot", line = 2.5)

# Display selected coefficients for the lambda.min model in Lasso
coef(lasso_model, s = lasso_lambda_min)

# Evaluation on test data for the final Lasso model
X_test <- model.matrix(emergency ~ . - 1, data = test)
Y_test <- test$emergency
pred_final <- predict(lasso_model, X_test, s = lasso_lambda_min, type = "response")

# Predicted vs Actual Plot for Lasso model
plot(Y_test, pred_final, xlab = "Actual Emergency Visits", ylab = "Predicted Emergency Visits",
     main = "Predicted vs Actual Emergency Visits (Lasso Model)")
abline(0, 1, col = "red")
```

```{r}
# Calculate log-likelihood on the test data for the selected model
log_likelihood <- sum(Y_test * log(pred_final) - pred_final - log(factorial(Y_test)))
log_likelihood
```

# Here’s a summary of the key results:

### Model Selection and Fitting:
  - A Negative Binomial model was fit to the training data to model emergency room visits, as instructed in the exercise.
  - Additionally, a Lasso regression with cross-validation was performed as the variable selection method, using a Poisson distribution.

### Model Summaries and Selected Coefficients:
  - The Negative Binomial model summary indicated significant variables and was displayed in the output.
  - The Lasso model identified non-zero coefficients based on the lambda.min value.

### Evaluation on Test Data:
  - The log-likelihood for the Lasso model on the test data was calculated and found to be approximately -4725.585, which assesses the fit of the selected model.
  - A Predicted vs. Actual plot for emergency room visits was generated, showcasing the performance of the Lasso model on test data.

# END.