---
title: "Michael_Ghattas_WP9"
author: "Michael Ghattas"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(survival)
library(survminer)
library(tidyverse)
library(ggplot2)
```

# Exercise:

1. Fit a Cox proportional hazard model to the data set with the predictors sex, and wt.loss (as main effects).  
2. Interpret the regression coefficients for the model.  
3. Predict the survival of a female who has wt.loss = 20.  
4. Assess the model assumptions

```{r}
attach(lung)
lung <- lung |>
  filter(!is.na(ph.ecog)) |>
  mutate(
    ph.ecog = factor(
      ph.ecog,
      levels = c(0, 1, 2, 3, 4),
      labels = c("Asymptomatic", "Ambulatory", "In bed <50%", "In bed >50%", "Bedridden")),
    ph.ecog = fct_drop(ph.ecog),
    ph.ecog = fct_lump_n(ph.ecog, n = 2, other_level = "Bedridden"),
    sex = factor(sex, levels = c(1, 2), labels = c("Male", "Female")),
    patient_id = row_number()
  )

head(lung)
```

### 1. Fit a Cox proportional hazard model to the data set with the predictors sex, and wt.loss (as main effects).

```{r ExerciseCox}
cox_lung <- coxph(Surv(time, status) ~ sex + wt.loss, data = lung)
summary(cox_lung)
```
### 2. Interpret the regression coefficients for the model.  

Interpretation:

  - The negative coefficient indicates that being female is associated with a lower hazard (risk of death) compared to being male.
  - The hazard ratio of 0.5865 means that females have approximately 41.35% lower hazard than males, holding wt.loss constant.
  - Since the p-value is less than 0.01, this effect is statistically significant.
  - The coefficient is close to zero, and the hazard ratio is approximately 1, suggesting no significant effect of wt.loss on the hazard.
  - The p-value is very high, indicating that variations in wt.loss are not significantly associated with the hazard rate in this model.
  
### 3. Predict the survival of a female who has wt.loss = 20. 

```{r}
# Define the new data point
new_data <- data.frame(sex = "Female", wt.loss = 20)

# Obtain the survival curve
surv_pred <- survfit(cox_lung, newdata = new_data)

# Display the survival probabilities at specific times
summary(surv_pred, times = c(100, 200, 300, 400, 500))
```

Interpretation:

  - Survival Probability at Time 100 Days: There's a 91.2% chance that the patient will survive beyond 100 days.
  - At 200 Days: The survival probability decreases to 78.3%.
  - At 300 Days: It further decreases to 65.1%.
  - At 400 Days: The probability is 52.1%.
  - At 500 Days: There's a 43.4% chance of surviving beyond 500 days.
  - The standard error and 95% confidence intervals provide information about the precision of these estimates.

### 4. Assess the model assumptions

```{r}
# Test proportional hazards assumption
ph_test <- cox.zph(cox_lung)
print(ph_test)
```
Interpretation:

  - Since the p-value is greater than 0.05, we fail to reject the null hypothesis that the effect of sex is proportional over time. Thus, the PH assumption holds for sex.
  - With a high p-value, there's no evidence against the PH assumption for wt.loss.
  - The global test also suggests that the PH assumption holds for the model overall.
  - The test results indicate that the PH assumption is not violated for either of the predictors or the model as a whole.

# END.