---
output:
  pdf_document: default
  html_document: default
---
title: "Michael_Ghattas_WP7"
author: "Michael Ghattas"
date: "`r Sys.Date()`"
output: pdf_document
---

# Environment Setup

```{r setup, include=FALSE}
# Set up the environment and suppress echo for cleaner output
knitr::opts_chunk$set(echo = TRUE)
# Load necessary libraries
library(mlbench)
library(glmnet)
```

# Load and Prepare Data

```{r}
# Load datasets: BreastCancer for classification and BostonHousing for regression
data("BreastCancer", package = "mlbench")
data("BostonHousing", package = "mlbench")

# Remove any rows with missing values and unnecessary ID column in BreastCancer data
BreastCancer <- na.omit(BreastCancer)[,-c(1)]
```

# Split BreastCancer Data into Train and Test Sets

```{r}
set.seed(1)

# Split data: 70% for training, 30% for testing
index <- sample(1:nrow(BreastCancer), 0.7 * nrow(BreastCancer), replace = FALSE)
train <- BreastCancer[index,]
test <- BreastCancer[-index,]

# Prepare matrices for ridge logistic regression
X.train <- model.matrix(Class ~ ., data = train)[,-1]  # Training predictors 
y.train <- as.numeric(train$Class) - 1  # Convert Class to binary numeric (0/1)

X.test <- model.matrix(Class ~ ., data = test)[,-1]   # Test predictors
y.test <- as.numeric(test$Class) - 1  # Convert Class to binary numeric (0/1)
```

# Ridge Logistic Regression with Cross-Validation on BreastCancer Data

```{r}
# Perform cross-validated ridge regression (alpha = 0) with logistic family
cvfit.ridge <- cv.glmnet(x = X.train, y = y.train, alpha = 0, family = 'binomial')

# Display lambda that minimizes cross-validation error and corresponding coefficients
cvfit.ridge$lambda.min
coef(cvfit.ridge, s = "lambda.min")
```

# Prepare BostonHousing Data for Lasso Regression

```{r}
# Set up design matrix for BostonHousing regression (exclude intercept) and response vector
X <- model.matrix(medv ~ ., data = BostonHousing)[,-1]
y <- BostonHousing$medv
```

# Lasso Regression with Cross-Validation on BostonHousing Data

```{r}
set.seed(202211) 

# Perform cross-validated lasso regression (alpha = 1)
cvfit.house.lasso <- cv.glmnet(x = X, y = y, alpha = 1)
# Display lambda that minimizes cross-validation error
cvfit.house.lasso$lambda.min

# Display the coefficients associated with lambda.min
coef(cvfit.house.lasso, s = "lambda.min")
```

